---
title: "OF analsis with DLCAnalyzer"
author: "Hannah Lapp"
date: "2/26/24"
output: pdf_document
params: 
  input_csv: "default.csv"
---
```{r,echo=FALSE,message=FALSE,warning=FALSE}
# Set so that long lines in R will be wrapped:
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```
## R Markdown

# Set up the workspace  
## Load libraries  
You may need to install the packages below if the are not already installed.  
```{r echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(sp)         #tested with v1.3-2
library(imputeTS)   #tested with v3.0
library(ggplot2)    #tested with v3.1.0
library(ggmap)      #tested with v3.0.0
library(data.table) #tested with v1.12.8
library(cowplot)    #tested with v0.9.4
library(corrplot)   #tested with v0.84
```
## Add DLC Analyzer functions  
We will need several functions from DLC Analyzer. You should download their code here: https://github.com/ETHZ-INS/DLCAnalyzer.
The specific file we need is in DLCAnalyzer-master/R/DLCAnalyzer_Function_Final.R. Move that file to you working directory.
Next, run the line below to import the functions from that file.
```{r}
source('./DLCAnalyzer_Functions_final.R')
```

# Import your dlc data  
Change the file name below to the deeplabcut csv output you want to analyze.
```{r tidy=TRUE, tidy.opts=list(width.cutoff=60)}
  data1 <- paste("Z:/hannah/automated_anxiety-main/Lab meeting examples/", 
                params$input_csv, sep ="")

#df <- read.csv(data)[,-1]
#rewrite_csv <- 
#write.csv(df, row.names=FALSE, file = data)

Tracking <- ReadDLCDataFromCSV(file = data1, fps = 30)
```

Plot histograms of the likelihood scores for a few key body points that will be used in the analysis. If the tracking is good, most points should be near 1 with only a few below .5.  

Nose:
```{r warning=FALSE, message=FALSE}
Tracking$data$nose %>% ggplot(aes(x = likelihood))+
  geom_histogram()+theme_bw()+
  ggtitle("Nose likelihood")

Tracking$data$top_head %>% ggplot(aes(x = likelihood))+
  geom_histogram()+theme_bw()+
  ggtitle("Top_head likelihood")

Tracking$data$back_3 %>% ggplot(aes(x = likelihood))+
  geom_histogram()+theme_bw()+
  ggtitle("Back_3 likelihood")
```


Similarly, we can find out how many frames have likelihood scores below .5.   

Top of the head:
```{r echo=FALSE}
print(paste("Number of frames top_head low likelihood:", (nrow(filter(Tracking$data$top_head, likelihood <.5)))))

print(paste("Number of frames back_3 low likelihood:", (nrow(filter(Tracking$data$back_3, likelihood <.5)))))

print(paste("Number of frames back_3 low likelihood:", (nrow(filter(Tracking$data$back_3, likelihood <.5)))))
```

We can also evaluate the tracking by plotting all the positions of a body parts across all frames. 
```{r echo=FALSE}
PlotPointData(Tracking, points = c("nose","top_head", "back_3"))
```
    
Assuming that the majority of points have high likelihood scores, we can proceed. The few points with low likelihood scores will get cleaned up later.  

Let's clean up points with likelihood scores below .5
```{r}
Tracking <- CleanTrackingData(Tracking, likelihoodcutoff = 0.5)
```

Check the plots to see if cleaning up improved tracking:  
```{r}
#check to see if cleaning up helped
PlotPointData(Tracking, points = c("nose","top_head", "back_3"))
```

Calibrate measurement units by converting from pixels to cm  
```{r}
Tracking <- CalibrateTrackingData(Tracking, method = "area",in.metric = 60*60, points = c("top_left_corner","top_right_corner","bottom_right_corner","bottom_left_corner"))
Tracking$px.to.cm
```

# Set the analysis zones  
```{r echo = FALSE}
Tracking <- AddOFTZones(Tracking, scale_center = 0.5,scale_periphery  = 0.6666 ,scale_corners = 0.34, points =c("top_left_corner","top_right_corner","bottom_right_corner","bottom_left_corner")) #center is set to half of arena (30 cm), periphery is set to 10cm
PlotZones(Tracking)
```

Convert names of points over to DLCAnalyzer names:  
```{r echo = FALSE}
tracking_points <- c( "nose",
                      "btwn_eyes",
                      "earl",
                      "earr",
                      "headcentre",
                      "neck",
                      "back_2",
                      "bodycentre",
                      "back_4",
                      "back_5",
                      "back_6",
                      "tailbase",
                      "left_side",
                      "right_side",
                      "tl",
                      "tr",
                      "bl",
                      "br",
                      "arena_center") 
names(Tracking$data) <- tracking_points
```

#Calculate speed and movement  
Using center of body
```{r}
Tracking <- CalculateMovement(Tracking, movement_cutoff = 5, integration_period = 5)
```


# Do some analysis!  
Create zone reports  
```{r}
Report_center <- ZoneReport(Tracking, point = "bodycentre", zones = "center")
Report_periphery <- ZoneReport(Tracking, point = "bodycentre", zones = "periphery")
Report_tl_corner <- ZoneReport(Tracking, point = "bodycentre", zones = c("corner.top_left_corner"))
Report_tr_corner <- ZoneReport(Tracking, point = "bodycentre", zones = c("corner.top_right_corner"))
Report_bl_corner <- ZoneReport(Tracking, point = "bodycentre", zones = c("corner.bottom_left_corner"))
Report_br_corner <- ZoneReport(Tracking, point = "bodycentre", zones = c("corner.bottom_right_corner"))
                                                                      
Report_arena <- ZoneReport(Tracking, point = "bodycentre", zones = "arena")
t(data.frame(Report_arena))
t(data.frame(Report_center))
t(data.frame(Report_periphery))
t(data.frame(Report_tl_corner))
t(data.frame(Report_tr_corner))
t(data.frame(Report_bl_corner))
t(data.frame(Report_br_corner))
```
  
# Create plots  
Zone visits:  
```{r}
PlotZoneVisits(Tracking,point = c("bodycentre","nose","tailbase"))
```
  
Density plots:  
```{r echo = FALSE}
plots <- PlotDensityPaths(Tracking,points = c("bodycentre","nose","tailbase"))
plots <- AddZonesToPlots(plots,Tracking$zones)
plots$bodycentre
plots$nose
plots$tailbase
```
